Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-06-23 article.tex
fix

Under the assumption that random variables follow a joint Gaussian
distribution $\mathcal{N}(\mu; \Sigma)$ of mean vector $\mu$ and covariance
matrix $\Sigma$, $X_i$ and $X_j$ are conditionally independent with respect to
the other variables $X^{-i,j}$ if their \textit{partial correlation} statistic
is null~\citep{baba2004partial}, i.e., if
\begin{equation}
\rho_{X_i, X_j | X^{-i,j}} = -\frac{\Sigma^{-1}_{ij}}{\sqrt{\Sigma^{-1}_{ii} \Sigma^{-1}_{jj}}} = 0.
\end{equation}
Conversely, $X_i$ and $X_j$ are conditionally dependent if $\rho_{X_i, X_j |
X^{-i,j}} \neq 0$, where $\rho_{X_i, X_j | X^{-i,j}} \in [-1;1]$ can now be
interpreted as the degree of association between the two variables (i.e., the larger the
value, the stronger the interaction). In practice due to the finite size of
time series, noise and modeling assumptions, the partial correlation
$\rho_{X_i, X_j | X^{-i,j}}$ will never be zero. If instead of a degree
of association, one would want to infer the underlying graph a statistical
test would be required to assert conditional independences. Note also that
the partial correlation
statistic is symmetric (i.e., $\rho_{X_i, X_j | X^{-i,j}} = \rho_{X_j, X_i |
X^{-j,i}}$) and therefore does not account for the direction of the
interaction. Together, the partial correlation statistics correspond to the
inverse $\Sigma^{-1}$ of the covariance matrix, also known as the concentration
or precision matrix. Computationally, they can be estimated efficiently in
$O(M^3)$ (instead of $O(p^3)$) from the $M$ first principal components
extracted through PCA~\citep{bishop2006pattern} on the processed data. In our
experiments, $M$ is set to $0.8 p$, thereby allowing us to further filter the
the data by removing the contribution of principal components
likely due to noise.

>>>>

Under the assumption that random variables follow a joint Gaussian
distribution $\mathcal{N}(\mu; \Sigma)$ of mean vector $\mu$ and covariance
matrix $\Sigma$, $X_i$ and $X_j$ are conditionally independent with respect to
the other variables $X^{-i,j}$ if their \textit{partial correlation} statistic
is null~\citep{baba2004partial}, i.e., if
\begin{equation}
\rho_{X_i, X_j | X^{-i,j}} = -\frac{\Sigma^{-1}_{ij}}{\sqrt{\Sigma^{-1}_{ii} \Sigma^{-1}_{jj}}} = 0.
\end{equation}
Conversely, $X_i$ and $X_j$ are conditionally dependent if $\rho_{X_i, X_j |
X^{-i,j}} \neq 0$, where $\rho_{X_i, X_j | X^{-i,j}} \in [-1;1]$ can now be
interpreted as the degree of association between the two variables (i.e., the larger the
value, the stronger the interaction). In practice due to the finite size of
time series, noise and modeling assumptions, the partial correlation
$\rho_{X_i, X_j | X^{-i,j}}$ will never be zero. If instead of a degree
of association, one would want to infer the underlying graph a statistical
test would be required to assert conditional independences. Note also that
the partial correlation
statistic is symmetric (i.e., $\rho_{X_i, X_j | X^{-i,j}} = \rho_{X_j, X_i |
X^{-j,i}}$) and therefore does not account for the direction of the
interaction. Together, the partial correlation statistics correspond to the
inverse $\Sigma^{-1}$ of the covariance matrix, also known as the concentration
or precision matrix. Computationally, they can be estimated efficiently in
$O(M^3)$ (instead of $O(p^3)$) from the $M$ first principal components
extracted through PCA~\citep{bishop2006pattern} on the processed data. In our
experiments, $M$ is set to $0.8 p$, thereby allowing us to further filter
the data by removing the contribution of principal components
likely due to noise.
